A statistical mechanics approach to the simulation of extreme fluctuations of the time-averaged drag acting on an obstacle immersed in a turbulent flow

### 2019-01-07 draft_p2.tex
Ré-itération générale du papier

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
  \caption{Typical vorticity field in the computational domain. The square obstacle is represented by the gray area in the center of the channel.}
\end{figure}
%% Description de la géométrie
We consider the interaction of a two dimensional turbulent channel flow with a square cylinder, illustrated in figure~\ref{fig:computational_domain}.
At the inlet of the channel, a constant parabolic velocity profile is imposed.
We note $u_0$ the imposed fluid velocity at the inlet, at the center of the channel.
Shortly after the inlet, the flow interacts with a grid, thus leading to turbulence.
Generated eddies are advected downstream by the mean-flow and grow in size along the way---a consequence of the inverse energy cascade taking place in two-dimensional turbulent flows.
A square obstacle is mounted in the channel.
Its distance from the grid is such that the typical size of the eddies it interacts with is close to the siye of the obstacle itself.
Fluid structures are then advected further downstream where they are damped in a region of increasing viscosity, a so-called \textit{sponge layer}~\cite{ref1,ref2}.
The outlet is modelled as an open boundary, where a zero momentum flux along the mean-flow direction in imposed.

@@

%% Descritpion d'une fluctuations de trainée typique
The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts exterted on the obstacle.
In particular, the \textit{drag force} is defined as the result of both pressure and viscous efforts, projected along the mean-flow direction, that is
\begin{equation}
  \label{eq:drag_definition}
  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ the identiy matrix, $\boldsymbol{\tau}$ the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.
to the formation of boundary layers at the top and bottom boundaries of the obstacle.
Figure~\ref{fig:typical_drag_signal} illustrates the time evoluction of the drag acting on the square obstacle.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
  \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}. The correlation time $\tau_c$ is defined further in this text and correspond to the typical fluctuation timescale of the drag.}
  \label{fig:typical_drag_signal}
\end{figure}
The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients fields in the vicinity of the obstacle.
Furthermore, we observed that the contribution of viscous stresses to the overall drag is three order of magnitudes smaller than the contrbution of pressure.
This a consequence of the blunt shape of the ostacle, that results in an important pressure difference between the forebody (the upstream boundary) and the base (the downstream boundary).
As a result, the drag acting on the square cylinder is very well approximated by
\begin{equation}
  \label{eq:drag_approx}
  f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
where $p_{fb}(t)$ amd $p_{base}(t)$ denote the pressure integrated over the forebody and the base, respectively.

The typical timescale of variation of the drag is therefore corresponds to the typical timescale of variation of the pressure in the vicinity of the obstale.
It can be estimated by the \textit{turnover time}
\begin{equation}
  \label{eq:turnover_time}
  \tau_0 = \frac{R}{U}
\end{equation}
where $R$ is the diamenter of the cylinder and $U$ the mean-flow velocity.

%% The phenomenology of typical drag fluctuations

Figure~\ref{fig:typical_drag_fluctuation} illustrate the evolution of the vorticity field in the vicinity of the obstacle.
Over a few turnover times, boundary layers form along the top and bottom boundaries.
The generated vorticity turns into eddies that are advected downstream by the mean flow.
As the result, the vorticity generated along the top and bottom boundary layers have a very small impact
on the the pressure field in the vicinity of the base of the obstacle.
As result, typical fluctuations of the drag are mostly due to the fluctating pressure in the vicinity of the upstream side of the obstacle.
Such fluctuating pressure directly orginates from the incoming turbulent flow.
In the following section~\ref{sec:direct_sampling}, we illustrate very different dynamics for \textit{extreme} fluctuations of the drag acting on the obstacle.
In particular, we show that extreme fluctuations primarily result from fluctuations of the base pressure, whereas forebody pressure fluctuations play very little role.

@@

In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of a very long timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ with $T_{tot} \approx 4\times 10^{6}\tau_0$.
Two comments are in order:
\begin{itemize}
\item The probabilty is skewed towards positive fluctuations. As an indication, the skewness coefficient is ??.
\item The tails of the distribution are well described by an exponential \ac{pdf}, \textit{i.e.} $\mathbb{P}(f_d=f) \propto e^{-\alpha f}\mathrm{d}f$.
\end{itemize}
The interaction of the flow is responsible for this particular shape of the \ac{pdf}.
This can be seen by considering the flow \textit{without} obstacle, and computing the \ac{pdf} of the drag acting on a control surface corresponding to the surface of the obstacle.
This experiment is depicted in figure~\ref{fig:no_square}.
One can see in figure~\ref{fig:pdf_drag} that the \ac{pdf} of the drag process exhibits a very different shape when no obstacle interacts with the flow.
In particular, it is symmetric and does not display exponential tails.
The exponential behaviour of the \ac{pdf} describing rare events plays an important role in the description of the phenomenology of extreme fluctuations of the \textit{time-averaged} drag.
This topic is further discussed in section~\ref{sec:avg_drag}.

Figure~\ref{fig:correlation_function} displays the autocorrelation function for the drag process, defined as 
\begin{equation}
  \label{eq:correlation_def}
  C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
\end{equation}
Time correlations decay exponentially.
As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.
From figure~\ref{fig:correlation_function}, we find that $\tau_c \approx 5\tau_0$.
From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.

@@

\paragraph*{Paragraphe d'introduction: les motivations d'un étude directe ?}
Rare event algorithms relies on the simulation of an ensemble of trajectories.
Members of the ensemble can be duplicated or removed from the ensemble, based on the evaluation of a \textit{score function}.
A score function is an obervable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achived by the trajectory.
In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observalbe over the trajectory.
However, for high-dimensional dynamics with a very complex phase space, the simple distance to the target event is likely to be a very poor criterion for selecting rare trajectories.
\textbf{[...]}

@@

We performed the simulation of the flow dynmaics described in section~\ref{sec:test_flow} over $T_{tot}=10^6\tau_c$.
The current value of the drag $f_d(t)$ was recorded at each time step, thus yielding a timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
As a post-processing step, we could identify the particular drag fluctuations with an amplitude greater than a fixed threshold amplitude $a$.
The threshold $a$ was then chosen so that the typical timescale of occurence of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
This typical timescale of occurence is referred to as the \textit{return time} and can be computed as a function of the fluctuation amplitude on the basis of the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$~\cite{lestang2018}.
Figure~\ref{return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard devaition of the drag process.
We denote by $(t^{\star}, f_d^{\star})$ an fluctation of amplitude $f_d^{\star}$ occuring at time $t^{\star}$.
On the basis of the drag timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we extracted the set of extreme fluctations
\begin{equation}
  \{(t^{\star}, f_d^{\star}) | f_d^{\star} \geq a\} \quad \mbox{with} \quad a=7.6\sigma.
\end{equation}
As a result, we extracted $104$ independent extreme drag fluctuations with a return time greater than $10^4\tau_c$.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{return_time_drag}
  \caption{Drag fluctuation amplitude as a function the correponding return time. The return time $r(a)$ is the typical timescale of occurence of a fluctuation of amplitude larger than $a$.}
  \label{fig:return_time_instant}
\end{figure}
\subsection{Instantaneous drag}
\label{sec:instantaneous_drag}

\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
\label{sec:forebody_and_base_contribution}
It was illustrated in section~\ref{sec:test_flow} that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations that are caused by the updtream turbulent flow.
In the following we investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuatons, in the case of the $104$ extreme drag events ${(t^{\star}, f_d^{\star})_n}_{1\leq n \leq 104}$ extracted from the timeseries.
For a given event , we consider the effective fluctuation, \textit{i.e.} the difference from the average value $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
it can be further decomposed into 
\begin{equation}
  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
\end{equation}
where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denotes the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.  
\begin{figure}
  \centering
  \label{fig:contribution_forebody_and_base_pressure}
\end{figure}
It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the drag fluctuation.
By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinderare dominated by the variation of the pressure in the vicinity of the base of the obstacle.
In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation relatively to the foreboy pressure fluctuation.

\subsubsection{Dynamical apsects of extreme drag fluctations}
\label{sec:dynamical_aspects}

@@

We now describe the dynamics associated with some of the extreme drag fluctuations sampled in the contrl timeseries.
Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events with the larger fluctuation in the whole timeseries.
Remarkably, all four events display very similar flow configurations.
In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels can reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
Such vorticity is produced through viscous shear along a boundary of the obstacle tangential to the mean flow direction.
By contrast with typical configurations, we observe that this vorticity is not directly advected by the mean flow, but instead constrained in the vicinity of the base of the obstacle.
Indeed, in each of the four examples displayed in figure~\ref{fig:top_four_events}, a large vortical region is located at roughly one obstacle length of the base.
As illustrated by the streamlines displayed in figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.

Figure~\ref{fig:dynamics} illustrates the dynamics leading to the extreme drag fluctuation corresponding to figure~\ref{fig:top_four_events_a}.
Figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} show that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy, advected by the flow above the obstacle.
The interaction of this eddy with the top boundary layer eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
This can be seen see in figure~\ref{fig:dynamics_k}.
Although this vorticity is too far from the base to directly affect the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle, as can be observed in figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.

Figure~\ref{fig:top_four_events_timeseries} shows the drag timeseries corresponding to the four events featured in figure~\ref{fig:top_four_events}.
The timeseries are centered on the extreme event.
It illustrates that these fluctuations are short-lived, as their temporel extension is rouhly one correlation time unit.
This is the time it takes for the block eddy to form and be advected away by the mean flow, thus frreeing the vorticity concentrated in the vicinity of the base of the obstacle.

Among the $104$ sampled events, $80\%$ can be related to very similar dynamics as the one described in the previous paragraph. In order to identify such events, we monitored the averaged shear along the top or bottom booundary:
\begin{equation}
  \label{eq:avg_shear_def}
  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x}
\end{equation}
where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary, depending on the event.
Figure~\ref{fig:type_1_and_type_2_a} displays the label of each sampled events, sorted according to the amplitude of the corresponding fluctuation.
It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the hghest fluctuation amplitude.
Figure~\ref{fig:type_1_and_type_2_a} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of these events.
For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
As illustrated in figure~\ref{fig:extremes_timeseries}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
Correspondingly, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
Interestingly, these excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
This confirms that the increase of $\overline{\gamma}$ acts as a precursor for extreme drag fluctuations.

The remaining $20\%$ of the sampled events correspond to slightly different dynmaics.
In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
This viscous shear is induced by a large vortex detached from the obstacle.

@@

% Motivation
In the previous section, we observed that rare events corresponding to extremely high values of the drag acting on the square obstacle have a lifetime of about $2\tau_c$.
In many applications, this duration is much smaller than the timescale of interest.
This is for instance the case when considering the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.

In such cases, a relevant observable is the \textit{time-averaged} drag
\begin{equation}
  \label{eq:def_time_averaged_drag}
  F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
\end{equation}
where $f_d$ denotes the instantaneous drag and $T\leq \tau_c$ a timescale of interest.
In the following we consider the case where $T \gg \tau_c$, typically $T=10\tau_c$.
In this context, a fluctuation of the average $F_T(t)$ is relatex to roughly $T / 2\tau_c$ independant fluctuations of $f_d$ over the time interval $[t;t+T]$.
What is the phenomenology leading to extreme values of $F_T(t)$ ?
Do an exceptionally large value of the average drag result from a single excpetionally large value of the instantaneous drag (case (1)) ?
Or from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
From the timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$ we deduced the timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$ and identified extreme positive fluctuations for which the drag exceeds a fixed threshold $a$ in the same way as described in section~\ref{sec:extreme_extraction}.
Setting $a=5.2\sigma _T$ leads to $84$ independant events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustration_vorticity_tf2}
  \caption{Typical vorticity field in the computational domain. The square obstacle is represented by the gray area in the center of the channel.}
\end{figure}
%% Description de la géométrie
We consider the interaction of a two dimensional turbulent channel flow with a square cylinder, illustrated in figure~\ref{fig:computational_domain}.
A constant parabolic velocity profile is imposed at the inlet of the channel and turbulence is
generated by means of a grid.
A square obstacle is mounted at the center of the channel.
Eddies are created in the vicinity of the grid with a typical size that correspond to the step of the grid.
As they are advected downstream, eddies interact with each other and grow in size, until they interact with the obstacle.
The distance separating the obstacle from the grid is such that the typical size of the eddies it interacts with is close to the size of the obstacle itself.
The flow in the vicinity of the outlet is regularized by means of a \textit{sponge layer}~\cite{ref1,ref2}, \textit{i.e.} a region where the viscosity is progressivaly increased.
Such sponge layer facilitates the implementation of an open boundary at the outlet, where a zero momentum flux along the mean-flow direction in imposed.

@@

%% Descritpion d'une fluctuations de trainée typique
The interaction of the incoming turbulent flow with the square obstacle results in fluctuating mechanical efforts.
In particular, the \textit{drag force} is defined as the result of both pressure and viscous constraints, projected along the mean-flow direction.
It reads
\begin{equation}
  \label{eq:drag_definition}
  f_d = \int_{\mathcal{S}}(-p(\boldsymbol{x})\boldsymbol{I} + \boldsymbol{\tau}) \mathrm{d}\boldsymbol{\mathcal{S}} \cdot \boldsymbol{e}_x,
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle, $p(\boldsymbol{x})$ is the pressure acting at location $\boldsymbol{x}$, $\boldsymbol{I}$ is the identity matrix, $\boldsymbol{\tau}$ denotes the viscous stress tensor and $\boldsymbol{e}_x$ is the unit vector pointing in the mean-flow direction.

%% Viscous effects are negligible
The value of the drag $f_d$ mainly depends on the local pressure and velocity gradients in the vicinity of the obstacle.
Furthermore, because of the blunt shape of the obstacle, the contribution of viscous constraints to the overall drag is negligage with respect to pressure constraints.
As a result, the drag mainly originates from the pressure difference between the \textit{forebody}---the upstream boundary---and the \textit{base}---the downstream boundary.
Consequently, the drag acting on the square cylinder is very well approximated by
\begin{equation}
  \label{eq:drag_approx}
  f_d(t) = p_{fb}(t) - p_{b}(t)
\end{equation}
where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base, respectively.

%% The phenomenology of typical drag fluctuations
The typical timescale of variation of the drag therefore corresponds to the timescale of pressure fluctuations in the vicinity of the obstacle.
An estimate of such timescale is the \textit{turnover time}
\begin{equation}
  \label{eq:turnover_time}
  \tau_0 = \frac{R}{U},
\end{equation}
where $R$ is the diameter of the cylinder and $U$ the mean-flow velocity.
Figure~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle over a few thousands turnover times.
The evolution over a few turnover times of the vorticity field around the obstacle is displayed in figure~\ref{fig:typical_vorticity}.
One can see that the interaction of the incoming flow with the square obstacle primarily results in boundary layers along the top and bottom boundaries.
However, the vorticity corresponding to these boundary layers only have a small impact on the the pressure field in the vicinity of the obstacle, because it is rapidly advected dowmstream by the mean-flow.
As a consequence, the base pressure varies much less than the forebody pressure, and typical drag fluctuations mostly result from the fluctuating pressure at the front of the obstacle.

In the following section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a smaller role.

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{illustration_signal_drag_1}
  \caption{Typical temporal evolution of the drag acting on the square cylinder pictured in figure~\ref{fig:computational_domain}. The correlation time $\tau_c$ is defined further in this text and correspond to the typical fluctuation timescale of the drag.}
  \label{fig:typical_drag_signal}
\end{figure}

@@

In the following we consider the drag force acting on the obstacle as a scalar random process $f_d(t)$.
In order to estimate its statistical properties, we performed a simulation of the dynamics over a duration $T_{tot} = 4\times 10^6\tau_c$.
In the following of this paper we refer to this long simulation as the \textit{control run}.
Figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} for the drag process, computed on the basis of the control timeseries $\{f_d(t)\}_{0\leq t \leq T_{tot}}$ obtained from the control run.
One can see that the tails of this distribution, that describe the probability of very rare events, can be very well modelled by an exponential \ac{pdf} of the form $p(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
In addition, figure~\ref{fig:pdf_drag} displays an estimate of the \ac{pdf} of the process describing the drag acting on a control surface equivalent to the surface of  the obstacle, in a flow \textit{without obstacle}.
This different setup is illustrated in figure~\ref{fig:no_square}.
In this case, the corresponding \ac{pdf} estimate for the drag process is quasi-symmetric and, more importantly, does not display exponential tails.
We conclude that the interation of the flow with the square obstacle is responsible for the exponential tails of the \ac{pdf} of the drag process.
In section~\ref{sec:time_avg}, we relate this very particular shape of the \ac{pdf} describing extreme drag fluctuations to the phenomenology of extreme fluctuations of the time-average of the drag process.

Lastly, figure~\ref{fig:correlation_function} illustrates that the drag process has exponentially decaying time correlations.
In the following, we call \textit{correlation time}, the typical time it takes for the drag process to decorrelate.
We note it $\tau_c$.
From figure~\ref{fig:correlation_function}, one can see that $\tau_c \approx 4\tau_0$.
This means that the drag decorrelates after only a few number of eddies interacted with the square obstacle.

% \begin{equation}
%   \label{eq:correlation_def}
%   C(\tau) = \frac{\mathbb{E}[f_d(t)f_d(t+\tau)] - \mathbb{E}^2[f_d]}{\sigma ^2} \quad \mbox{with} \quad \sigma^2 = \mathbb{f_d^2} - \mathbb{f_d}^2.
% \end{equation}

% As a consequence, temporal correlations in the drag process have a finite life time $\tau_c$---the \textit{correlation time}.

% From the point of view of the drag process, the system lost its memory after only a few eddies interacted with the obstacle.
% This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.

@@

\paragraph*{Paragraphe d'introduction: les motivations d'un étude directe ?}
%% TODO: Introduire la notion de temps de retour
Rare event algorithms relies on the simulation of an ensemble of trajectories.
Members of the ensemble can be duplicated or removed from the ensemble, based on the evaluation of a \textit{score function}.
A score function is an observable of the trajectories that can quantify how interesting is a trajectory with respect to the realisation of a rare event.
For instance, if one is interested in a rare transition between two region of phase space, a possible choice for the score function is the minimal distance to the target set achieved by the trajectory.
In cases where one is interested in computing rare trajectories over which a given observable reaches a fixed threshold, an obvious score function is the maximum value of the observable over the trajectory.
However, for high-dimensional dynamics with a very complex phase space, the simple distance to the target event is likely to be a very poor criterion for selecting rare trajectories.
\textbf{[...]}

@@

On the basis of the control timeseries $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$, we tracked every drag fluctuation with an amplitude greater than a fixed threshold amplitude $a$.
The value of this threshold was chosen so that the return time of fluctuations with an amplitude larger than $a$ is $T_{tot}/100$.
Figure~\ref{return_time_instant} displays the fluctuation amplitude $a$ as a function of the return time.
According to Figure~\ref{return_time_instant}, the threshold amplitude $a$ is set to $7.6\sigma$ with $\sigma$ the standard deviation of the drag process.
We denote by $(t^{\star}, f_d^{\star})$ a fluctuation of peak amplitude $f_d^{\star}$ occurring at time $t^{\star}$, as illustrated in figure~\ref{fig:extraction}.
From the control timeseries, we identified $104$ independent extreme drag fluctuations whith an amplitude greater than $7.6\sigma$, \textit{i.e.} with a return time greater than $10^4\tau_c$.
In the following sections~\ref{sec:forebody_and_base_contribution} and~\ref{sec:instantaneous_drag}, we describe the phenomenology of extreme drag fluctuations on the basis of this set of
events.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{return_time_drag}
  \caption{Drag fluctuation amplitude as a function the corresponding return time. The return time $r(a)$ is the typical timescale of occurrence of a fluctuation of amplitude larger than $a$.}
  \label{fig:return_time_instant}
\end{figure}
\subsection{Instantaneous drag}
\label{sec:instantaneous_drag}

\subsubsection{Contribution of forebody and base pressure fluctuations to the overall drag fluctuation}
\label{sec:forebody_and_base_contribution}
In section~\ref{sec:test_flow}, we pointed out that typical drag fluctuations mainly originate from fluctuations of the forebody pressure, \textit{i.e.} pressure fluctuations resulting from the upstream turbulent flow.
We now investigate the relative contribution of forebody and base pressure fluctuations to the overall drag fluctuations, in the case of the sampled \textit{extreme} drag fluctuations.
Let $(t^{\star}, f_d^{\star})$ be an extreme drag event.
In the following we consider the effective fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \langle f_d \rangle$.
It can be further decomposed into 
\begin{equation}
  \tilde{f}_d^{\star} = \tilde{p}_{fb}^{\star} - \tilde{p}_{base}^{\star},
\end{equation}
where $\tilde{p}_{fb}^{\star}$ and $\tilde{p}_{base}^{\star}$ denote the fluctuation of the forebody pressure and base pressure, respectively, that is $\tilde{p}_{base}^{\star} = p_{base}^{\star} - \langle p_{base} \rangle$ .
Figure~\ref{fig:contribution_forebody_and_base_pressure} shows the relative contributions $\tilde{p}_{base}^{\star}/\tilde{f}_d^{\star}$ and $\tilde{p}_{fb}^{\star}/\tilde{f}_d^{\star}$ to the overall drag fluctuation for the base and forebody pressure fluctuations, respectively.  
\begin{figure}
  \centering
  \label{fig:contribution_forebody_and_base_pressure}
\end{figure}
It shows that among the 104 sampled extreme events, the base pressure fluctuation typically contributes for $80\%$ of the overall drag fluctuation.
By contrast with typical drag fluctuations, extreme fluctuations of the drag acting on the square cylinder are dominated by the variation of the pressure in the vicinity of the base of the obstacle.
In addition, figure~\ref{fig:contribution_forebody_and_base_pressure} hints that, the larger the fluctuation, the more important is the contribution of the base pressure fluctuation, relatively to the forebody pressure fluctuation.

\subsubsection{Dynamical aspects of extreme drag fluctuations}
\label{sec:dynamical_aspects}

@@

We now turn to a description of the flow dynamics for some of the extreme drag fluctuations, sampled in the control timeseries.
Figure~\ref{fig:top_four_events} displays the vorticity field at $t=t^{\star}$ for the four events corresponding to the largest fluctuation amplitudes.
One can see that they correspond to very similar flow configurations.
In each case, a strong vorticity region is located in the vicinity of the base of the obstacle, where vorticity levels typically reach twice as much as the typical vorticity fluctuations displayed in figure~\ref{fig:typical_event}.
Such strong vorticity is produced through viscous shear along either the bottom or top boundary of the obstacle.

By contrast with typical configurations, we observe that this vorticity is not directly advected by the mean flow, but instead constrained in the vicinity of the base of the obstacle.
Indeed, in each of the four examples displayed in figure~\ref{fig:top_four_events}, a large vortical region is located at roughly one obstacle length of the base.
As illustrated by figure~\ref{fig:top_four_events_streamlines}, vorticity generated alongside the boundary has no other choice but to concentrate close to the base of the obstacle.

Figure~\ref{fig:top_four_events_timeseries} displays the drag timeseries corresponding to the four events illustrated in figure~\ref{fig:top_four_events}.
It illustrates that the lifetime of such fluctuations temporal extension is roughly one correlation time unit $\tau_c$.
It is the time it takes for the block eddy to form and be advected away by the mean flow, thus freeing the vorticity concentrated in the vicinity of the base of the obstacle.
This can be seen in figure~\ref{fig:dynamics}, which illustrate the vorticty dynamics from $t^{\star}-\tau_c/2$ to $t^{\star}$ on a particular example of an extreme drag fluctuation. 
On can see in figures~\ref{fig:dynamics_a} to~\ref{fig:dynamics_g} that, from roughly $t^{\star}-\tau_c/2$ to $t^{\star}-\tau_c/3$, the top boundary layer is perturbed by an incoming eddy.
This interaction eventually leads to the formation of a region of negative vorticity at roughly one square length of the base of the obstacle.
This can be seen see in figure~\ref{fig:dynamics_k}.
Although this vorticity is too far from the base to directly affect the base pressure, and thus the drag, it constrains the positive vorticity generated alongside the bottom boundary of the obstacle---see figures~\ref{fig:dynamics_k} to~\ref{fig:dynamics_p}.

We found that $80\%$ of the extreme events sampled from the control timeseries can be related to very similar dynamics.
In order to identify such events, we monitored the averaged shear along the top or bottom boundary
\begin{equation}
  \label{eq:avg_shear_def}
  \overline{\gamma} = \frac{1}{L} \int_{\mathcal{S}_b} \frac{\partial u_x(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
\end{equation}
where $L$ denotes the diameter of the cylinder, $u_x$ the longitudinal component of the velocity field and $\mathcal{S}_b$ the surface of either the top or the bottom boundary.
% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
Figure~\ref{fig:gamma_asof_fd} shows $\overline{\gamma}$ as a function of the instantaneous drag $f_d$, for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$, for each of the sampled events that follow the previously described dynamics..
For $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths concentrate in the region describing typical values for both $\overline{\gamma}$ and $f_d$.
As illustrated in figure~\ref{fig:extremes_timeseries}, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$.
As a consequence, paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does.
This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.

The remaining $20\%$ of the sampled events correspond to slightly different dynamics.
In this case the vorticity responsible for the base pressure drop is not created along the top or bottom boundary, but directly through viscous shear alongside the base boundary of the obstacle.
This viscous shear is induced by a large vortex detached from the obstacle.
For a more detailed description of such dynamics, see the chapter 3 of ~\ref{ref:these_lestang}.

@@

% Motivation
In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $tau_c$.
However, In many applications, this duration is much smaller than the timescale of interest.
Consider for insatnce the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.

%% Definition de la trainée intégrée
In such cases, a relevant observable is the \textit{time-averaged} drag
\begin{equation}
  \label{eq:def_time_averaged_drag}
  F_T(t) = \frac{1}{T}\int_t^{t+T} f_d(t) \mathrm{d}t,
\end{equation}
where $f_d$ denotes the instantaneous drag and $T\leq \tau_c$ a timescale of interest.
In the following we consider the case where $T \gg \tau_c$, typically $T=10\tau_c$.
In this context, a fluctuation of the average $F_T(t)$ is related to roughly $T / \tau_c$ independent fluctuations of the \textit{instantaneous} drag $f_d$, over the time interval $[t;t+T]$.

%% Quelle est la phénoménologie de la trainée intégrée ?
What is the phenomenology leading to extreme values of $F_T(t)$ ?
Do an exceptionally large value of the average drag result from a single exceptionally large value of the instantaneous drag (case (1)) ?
Or instead from a succession of rather typical fluctuations, however all in the same direction (case (2)) ?
From the control timeseries, see section~\ref{sec:pdfs}, one can compute a control timeseries $\{F_T(t)\}_{0 \leq t \leq T_{tot}-T}$.
It is then possible to identify extreme positive fluctuations for which the time-averaged $F_T$ drag exceeds a fixed threshold $a$, in the very same way as described in section~\ref{sec:extreme_extraction}.
Setting $a=5.2\sigma _T$ leads to $84$ independent events, with $\sigma_T$ the standard deviation of the random process describing the time evolution of $F_T$.
